---
title: "STAT545 hw04"
output: github_document
---

## Setting ups
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```

## Data Reshaping Prompts

### Activity #2

##### Make a tibble with one row per year and columns for life expectancy for two or more countries.

```{r}
lvsc <- gapminder %>%
  group_by(country) %>%
  filter(continent == "Americas" & pop[year == 2007] > 2e7) %>%
  select(year, country, lifeExp) %>%
  spread(key = "country", value = "lifeExp")
```

- Use `knitr::kable()` to make this table look pretty in your rendered homework.

```{r}
knitr::kable(lvsc)
```

- Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
lvsc %>%
  ggplot() +
  geom_point(aes(Canada, `United States`, color = year))
```


## Join Prompts

### Activity #1

##### Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins.

- One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city.

```{r}
ac <- gapminder %>%
  filter(continent == "Americas" & year == 2007) %>%
  select(-continent)
df1 <- tribble(
  ~country, ~`NATO membership`, ~`capital city`,
  "Canada", "Yes", "Ottawa",
  "United States", "Yes", "Washington",
  "Mexico", "No", "Mexico City",
  "Brazil", "No", "Brasilia",
  "Argentina", "No", "Buenos Aires"
)
```

Now we can explore `left_join`, `right_join`, `inner_join` and `full_join`.
```{r}
knitr::kable(left_join(ac, df1))
knitr::kable(right_join(ac, df1))
knitr::kable(inner_join(ac, df1))
knitr::kable(full_join(ac, df1))
```



- One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.

```{r}
conts <- gapminder %>%
  group_by(continent) %>%
  summarise(`country num` = length(country))
df2 <- tribble(
  ~continent, ~hemisphere,
  "Asia", "North",
  "Europe", "North",
  "Oceania", "South"
)
```

Again we can explore `left_join`, `right_join`, `inner_join` and `full_join` now.
```{r}
knitr::kable(left_join(conts, df2))
knitr::kable(right_join(conts, df2))
knitr::kable(inner_join(conts, df2))
knitr::kable(full_join(conts, df2))
```

### Activity #3

##### Explore `merge` and `match`

```{r}
knitr::kable(merge(df1, df2))
match(df1, df2)
match(ac$country, gapminder$country)
```